---
title: 'Data quality'
author: 'Jay Achar'
date: '`r Sys.time()`'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
suppressMessages(library(tidyverse))
library(here)
library(naniar)
library(patchwork)
library(flextable)
library(officer)
devtools::load_all()
```

```{r source-functions}
files <- list.files(here("R"))
walk(files,
     ~source(here("R", .x)))

# returns a list of constants
const <- constants()
```

```{r read-in-data, warning=FALSE}
raw <- read_data()
# creates a list of country and region specific data frames in
# wide and long format
dd <- prepare_data(raw)

nested <- dd$wide$country %>%
  nest(data = c(everything(), -year))
```

```{r define-functions }
describe_missingness <- function(df = dd$long$country,
                                 group = c("hbc", "all")) {

  group <- match.arg(group)

  stopifnot(is.data.frame(df))

  ad <- filter(df,
               age_group != "014",
               year >= 2014)

  if (group == "hbc") {
    ad <- filter(ad,
                 iso3 %in% const$high_burden)
  }

        countries_missing <- ad %>%
                group_by(iso3, country) %>%
                summarise(na = any(is.na(cases)),
                          .groups = "drop") %>%
                filter(na == TRUE) %>%
          select(iso3, country)

        # most recent year with no missing information for any HBC with missing data
        recent_no_na_year <- ad %>%
                filter(year <= 2019) %>%
                group_by(iso3, year) %>%
                summarise(not_na = all(!is.na(cases)),
                          .groups = "drop") %>%
                filter(not_na == TRUE) %>%
                group_by(iso3) %>%
                filter(year == max(year))

        annual_totals <- df %>%
                filter(year %in% recent_no_na_year$year,
                       year >= 2014,
                       age_group != "014") %>%
                group_by(year) %>%
                summarise(total_cases = sum(cases, na.rm = TRUE),
                          .groups = "drop")

        missing_count <- ad %>%
        group_by(iso3) %>%
        summarise(total_datapoints = n(),
                  missing_datapoints = sum(is.na(cases))) %>%
        filter(missing_datapoints > 0) %>%
        arrange(desc(missing_datapoints)) %>%
          select(everything(),
                 `Total data points` = total_datapoints,
                 `Total missing data points` = missing_datapoints)

        prop_contributed <- ad %>%
                filter(year %in% recent_no_na_year$year) %>%
                group_by(iso3, year) %>%
                summarise(cases = sum(cases, na.rm = TRUE),
                          .groups = "drop") %>%
                semi_join(recent_no_na_year,
                          by = c("iso3", "year")) %>%
                right_join(annual_totals, by = "year") %>%
                mutate(prop = round(cases / total_cases * 100, 2)) %>%
                arrange(desc(prop)) %>%
          select(-total_cases, annual_cases = cases, year) %>%
          right_join(countries_missing, by = "iso3") %>%
          select(iso3, country, everything()) %>%
          left_join(missing_count,
                    by = "iso3")


        return(
            prop_contributed
        )
}

```

## Missing values

Find all timeseries with missing values to all imputation.

```{r extract-missing-values}
# extract all time series with zero values
missing <- dd$long$country %>%
        filter(iso3 %in% const$high_burden) %>%
        filter(year >= 2014,
               year != 2020,
               age_group != "014") %>%
        mutate(age_group = factor(
                age_group,
                levels = c("04", "514", "15plus"),
                ordered = TRUE
        )) %>%
        select(iso3, year, sex, age_group, cases) %>%
        group_by(iso3, sex, age_group) %>%
        nest() %>%
        filter(map_lgl(data, function(d) {
                any(d$cases == 0 | is.na(d$cases))
        })) %>%
        unnest(cols = c(data)) %>%
        ungroup() %>%
        mutate(cases = ifelse(cases == 0,
                              NA_integer_,
                              cases))

# saveRDS(missing,
#         file = here::here("artefacts/missing.rds"))

```

## High Burden Countries { .tabset}

```{r hbc-calculations, include = FALSE}
hbc_missing <- describe_missingness(group = "hbc")
```

### Commentary

Out of 30 HBCs, `r nrow(hbc_missing)` report some missing data when
disaggregated by year, age group and sex. The list of countries with their annual total case
notifications (in a year with no missing data) is shown:

```{r missing-hbc-countries}

hbc_missing_table <- hbc_missing %>%
  select(-iso3, -starts_with("Total "),
         Country = country,
         `Reported year` = year,
         `Annual notifications` = annual_cases,
         `Annual total contribution (%)` = prop) 

knitr::kable(hbc_missing_table)

hbcdq1 <- flextable(hbc_missing_table) %>% 
  set_table_properties(layout = "autofit", width = 1) %>%
  fontsize(size = 9, part = "all") %>%
  theme_box()

```

### Missing by year

```{r hbc-missing-by-year, include = FALSE}
hbc_missing_by_year <- dd$long$country %>%
        filter(iso3 %in% const$high_burden,
               year >= 2014,
               age_group != "014") %>%
        group_by(year) %>%
        summarise(na = sum(is.na(cases)),
                  length = n(),
                  perc = round(na / length * 100, 2)) %>%
        ungroup() %>%
        mutate(cumsum = cumsum(na))
```

When disaggregated by sex, age group, HBC and year the total number of
data points should be `r sum(hbc_missing_by_year$length)`. However,
`r sum(hbc_missing_by_year$na)` data points are missing - here's a table
describing their distribution across study years:

```{r hbc-missing-year-table}
hbc_missing_by_year %>%
  select(-na, -length,
         Year = year,
         `Missing data points (%)` = perc,
         `Cumulative missing data points` = cumsum) %>%
  knitr::kable()
```

### Missing data

Review how many years each high burden country has reported data for each
age group:

```{r hbc-data-availability}
hbc <- dd$long$country %>%
  filter(iso3 %in% constants()$high_burden,
         year >= 2014,
         age_group %in% c("04", "514", "15plus")) %>%
  mutate(age_group = factor(age_group,
                            levels = c("04", "514", "15plus"),
                            ordered = TRUE)) %>%
  group_by(iso3, year, age_group) %>%
  summarise(all_cases = sum(cases))

hbc %>%
  ggplot(aes(x = year,
             y = iso3,
             fill = is.na(all_cases))) +
  geom_tile(color = "white",
            lwd = 0.5,
            linetype = 1) +
  scale_fill_discrete(type = c("#0a9396", "#ae2012"),
                      labels = c("True", "False")) +
  labs(y = NULL,
       x = NULL,
       fill = "Annual data reported?",
       title = "HBC data reporting by age-group") +
  facet_grid(~ age_group) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

Red tiles correspond to a year when a given country did not report both male and female
notifications for the age-group. If either male or female data is missing,
the tile is coloured red.

### Relative change by year

```{r hbc-reporting}

hbc %>%
  group_by(iso3, age_group) %>%
  mutate(rel_change = ifelse(all_cases > 1000,
                             all_cases / lag(all_cases, 1),
                             1)) %>%
  ggplot(
    aes(x = year,
        y = iso3,
        fill = log(rel_change))
  ) +
  geom_tile() +
  facet_grid(cols = vars(age_group)) +
  labs(x = NULL,
      y = NULL,
      fill = "Log percent change from previous year",
      title = "HBC relative change in annual case notifications by age-group") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

The log percent change from one year to the next is plotted by country and age group. High
positive numbers (representing large increases in notifications) and numbers less than zero
(decreases in notifications) are worth looking more carefully at.

### All cases

```{r hbc-quality}

hbc_plots <- hbc %>%
  group_by(iso3) %>%
  nest() %>%
  mutate(plots = map2(data, iso3, function(df, country) {
    ggplot(df, aes(x = year, y = all_cases, group = age_group)) +
      geom_line() +
      facet_wrap(facets = vars(age_group),
                 ncol = 3,
                 scales = "free_y") +
      theme_minimal() +
      labs(title = country) +
      scale_y_continuous(limits = c(0, max(df$all_cases) * 1.1))

  }))

hbc_plots$plots
```

## All Countries { .tabset }

```{r all-countries-xcalculations, include = FALSE}
all_countries_missing <- describe_missingness(group = "all")
```

### Commentary

Out of `r length(unique(dd$long$country$iso3))` countries,
`r nrow(all_countries_missing)` report some missing data when
disaggregated by year, age group and sex. The list of countries
with their annual total case notifications (in a year with no
missing data) is shown:

```{r missing-all-countries}

all_countries_missing_table <- all_countries_missing %>%
  select(iso3, -starts_with("Total "),
         Country = country,
         `Reported year` = year,
         `Annual notifications` = annual_cases,
         `Annual total contribution (%)` = prop)

knitr::kable(all_countries_missing_table %>% select(-iso3))

acdq1 <- all_countries_missing_table %>% 
  filter(!iso3 %in% const$high_burden) %>% 
  select(-iso3) %>% 
  slice_max(`Annual total contribution (%)`,
            n = 10) %>% 
  flextable() %>% 
  set_table_properties(layout = "autofit", width = 1) %>%
  fontsize(size = 9, part = "all") %>%
  theme_box()

```

### Country names for manuscript

```{r result = "asis"}
all_countries_missing %>%
  arrange(country) %>%
  pull(country) %>%
  paste0(collapse = ", ")

```

## Missing values {.tabset}

The tables below show the number and proportion of missing values
in each notification variable by year. Notice how the number of missing
values is almost the same in any given age-band. This is
explored further in the next section.

### Univariable missingness

> Summary: 0-4y and 5-14y variables have more missing values - upto 20%; no difference by sex. HBC have improved dramatically in recent years.

```{r missing-var-plot, warning=FALSE}
# select data variables
data_vars <- str_detect(names(dd$wide$country),
                        paste(c(const$age_groups, "year"), collapse = "|"))
gg_miss_var(dd$wide$country[, data_vars],
            facet = year, show_pct = TRUE)

```

### HBC univariable missingness

```{r hbc-missing-var-plot, warning=FALSE}
dd$wide$country %>%
  filter(iso3 %in% const$high_burden) %>%
  gg_miss_var(facet = year, show_pct = TRUE)

```

### Missingness interactions by year

> Countries tend ot report missing values either across all variables or 0-4y & 5-14y

```{r missing-interactions}
nested %>%
  mutate(interactions = map(data,
                            ~ gg_miss_upset(.x, nsets = 8)) %>%
           set_names(year)) %>%
  pull(interactions)
```

### Tables

> Included for completeness more than anything - includes absolute and relative missingness

```{r country-missing}
nested %>%
  mutate(country_missing = map(
    data,
    function(yr) {
      df <- data.frame(
        variable = names(yr),
        missing_n = map_dbl(yr, naniar::n_miss),
        missing_prop = round(map_dbl(yr, naniar::prop_miss), 3)
      )
      row.names(df) <- NULL
      return(df)
    }) %>%
      set_names(year)) %>%
  pull(country_missing)

```

## Age-band missingness

Below are plots summarising missingness within
each age-band. The first, larger plot shows the
association between sex (green points) and the distribution of
missing values (red points). This is an alternative way of visualising
the plots above.

The small plots below show countries where one sex
has missing data. Notice that there are few countries where
this is the case, and that the number of notifications in the
avaialble sex is low suggesting that the missingness is unlikely
to have altered analysis results.

These plots are not organised by year since there's relatively few
missing points to show.

> It's rare for missing values to only be reported in one sex

```{r country-missing-distribution}
all_missing_plot <- function(df,
                         age_group,
                         title) {


  male <- paste0("newrel_m", age_group)
  female <- paste0("newrel_f", age_group)

  ggplot(df,
         aes(x = .data[[male]],
             y = .data[[female]])) +
    geom_miss_point() +
    labs(title = title)

}

sex_missing_plot <- function(df,
                             age_group,
                             title) {

  male <- paste0("newrel_m", age_group)
  female <- paste0("newrel_f", age_group)

  f <- df %>%
    filter(is.na(.data[[male]]),
           ! is.na(.data[[female]])) %>%
    ggplot(aes(x = .data[["iso3"]],
             y = .data[[female]])) +
  geom_point() + coord_flip()

  m <- df %>%
    filter(is.na(.data[[female]]),
           ! is.na(.data[[male]])) %>%
    ggplot(aes(x = .data[["iso3"]],
             y = .data[[male]])) +
  geom_point() + coord_flip()

  list(f, m)
                             }


all_missing_plots <- function(age_group) {
  mp <- all_missing_plot(
    dd$wide$country,
    age_group,
    title = age_group
  )

  sp <- sex_missing_plot(
    dd$wide$country,
    age_group = age_group,
    title = age_group
  )

  mp / (sp[[1]] | sp[[2]])
}

map(const$age_groups,
    all_missing_plots)


```

## Comparing notifications {.tabset}

In this section, I've displayed the aggregated 0-14y
notifications by sex where either the 0-4y or 5-14y values
are missing.

The first table shows all countries sorted by the value of
the aggregate variable (0-14y).

The second table is restricted to the 30 high-burden countries.

> There are very small discrepancies between 0-14y notifications and 0-4y + 5-14y notifications across years and sexes. When restricted to the HBC, the discrepancy is minimal

### Female

#### All countries

```{r aggregate-notifications-female}

aggregated_female <- nested %>%
  mutate(tables = map(data, function(x) {
    total <- sum(x$newrel_f014, na.rm = TRUE)
    df <- x[, c("iso3", "newrel_f014", "newrel_f04", "newrel_f514")]
    df$sum <- df$newrel_f04 + df$newrel_f514
    df$diff <- df$newrel_f014 - df$sum
    df <- df[which(is.na(df$diff)), ]
    df <- df[which(!is.na(df$newrel_f014)), ]
    df$all_total_prop <- round(df$newrel_f014 / total, 2)
    df %>% arrange(desc(newrel_f014)) %>%
      select(iso3, newrel_f014, all_total_prop)

  }) %>% set_names(year)
  )

aggregated_female %>% pull(tables)
```

#### 30 High burden countries

```{r notificaiton-difference-hb-female}
aggregated_female %>%
  mutate(hb = map(tables, ~ filter(.x, iso3 %in% const$high_burden))) %>%
  pull(hb)

```

```{r}
# aggregate_female %>%
#   slice_head(n = 15) %>%
#   ggplot(aes(x = iso3,
#              y = newrel_f014)) +
#   geom_col() + coord_flip()
```

### Male

#### All countries

```{r aggregate-notifications-male}

aggregated_male <- nested %>%
  mutate(tables = map(data, function(x) {
    total <- sum(x$newrel_m014, na.rm = TRUE)
    df <- x[, c("iso3", "newrel_m014", "newrel_m04", "newrel_m514")]
    df$sum <- df$newrel_m04 + df$newrel_m514
    df$diff <- df$newrel_m014 - df$sum
    df <- df[which(is.na(df$diff)), ]
    df <- df[which(!is.na(df$newrel_m014)), ]
    df$all_total_prop <- round(df$newrel_m014 / total, 2)
    df %>% arrange(desc(newrel_m014)) %>%
      select(iso3, newrel_m014, all_total_prop)
  }) %>% set_names(year)
  )

aggregated_male %>% pull(tables)
```

#### 30 High burden countries

```{r notificaiton-difference-hb-male }
aggregated_male %>%
  mutate(hb = map(tables, ~ filter(.x, iso3 %in% const$high_burden))) %>%
  pull(hb)
```

```{r print-docx, eval=FALSE}
ps_landscape <- prop_section(page_size = page_size(orient = "landscape"))
ps_portrait <- prop_section(page_size = page_size(orient = "portrait"))

tables_doc <- read_docx() %>%
  body_add_par("HBC Missing country contribution data quality table",
               style = "heading 1") %>%
  body_add_flextable(value = hbcdq1) %>%
  body_add_break() %>%  
  body_add_par("Top 10 missing country contribution data quality table",
               style = "heading 1") %>%
  body_add_flextable(value = acdq1) %>%
  body_add_break() %>%  
  body_end_block_section(value = block_section(property = ps_landscape))

print(tables_doc, target = here::here("figures/data_quality_tables.docx"))


```

