---
title: 'Generate manuscript artefacts'
author: 'Jay Achar'
date: '`r Sys.time()`'
params:
  save_artefacts: false
  plot_type: 'png'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(forecast)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(patchwork)
library(flextable)
library(officer)
library(data.table)
library(metafor)
library(ggrepel)
library(ggpubr)
library(purrr)
library(wbmapdata) # devtools::install_github('petedodd/wbmapdata')
library(sf)

absspace <- function(x, ...) {
  format(abs(x), ..., big.mark = " ",
         scientific = FALSE,
         trim = TRUE)
}

lancet_palette <- c("#00458B",
                    "#ED2300",
                    "#43B440",
                    "#0199B4",
                    "#915E9F",
                    "#DF9043")

# returns a list of constants
const <- constants()

raw <- suppressWarnings(read_data())
dd <- prepare_data(raw, const = const)
clean <- prepare_long_data(dd, const = const)

preds <- readRDS(paste0(output_dir,
                        "/manuscript_predictions.rds"))


```

## Figures {.tabset .tabset-pills}

### Main figure 1

```{r main-figure-1 }
# create ts_df list
ts_list <- clean$region %>%
        group_by(location, year, age_group) %>%
        summarise(cases = sum(cases),
                  .groups = "drop") %>%
        group_by(age_group) %>%
        group_split()

forecast_list <- preds$age_region %>%
        select(-sd, region = location) %>%
        mutate(year = 2020) %>%
        group_by(age_group) %>%
        group_split()

ribbon_df <- purrr::map2(ts_list,
                  forecast_list,
     function(ts, forecast) {
       ts %>%
        filter(year == 2019) %>%
        mutate(lower95 = cases,
               upper95 = cases) %>%
        select(age_group,
               region = location,
               point = cases,
               lower95, upper95,
               year) %>%
        bind_rows(forecast)
     })

age_group_labels <- c("0-4 years",
                      "5-14 years",
                      "15+ years")

y_breaks <- list(
  c(0, 5000, 10000, 25000, 50000, 75000, 100000),
  c(0, 10000, 25000, 50000, 75000, 100000, 125000, 150000),
  c(0, 250000, 500000, 1000000, 1500000, 2000000, 2500000, 3000000)
)

mf1 <- purrr::pmap(list(ts_list,
     forecast_list,
     ribbon_df,
     age_group_labels,
     y_breaks),
     function(ts, forecast, ribbon, age_label, y_break) {
       ggplot(ts,
              aes(x = year)) +
         geom_line(aes(y = cases,
                       color = location)) +
         geom_point(aes(y = cases,
                        color = location),
                    size = 1) +
         geom_ribbon(
           data = ribbon,
           mapping = aes(
             ymin = lower95,
             ymax = upper95,
             group = region,
             fill = region
           ),
           alpha = 0.3
         ) +
         scale_color_discrete(guide = "none", type = lancet_palette) +
         scale_fill_discrete(guide = "none", type = lancet_palette) +
         scale_y_sqrt(
           expand = c(0, 0),
           label = absspace,
                      # limits = c(0, 100000),
                      breaks = y_break,
                      minor_breaks = NULL) +
         labs(x = "Year",
              y = "Cases notifications",
              fill = "WHO Region",
              title = age_label) +
         theme_minimal() +
         expand_limits(y = 0)

     }) %>%
  wrap_plots(ncol = 3) +
  scale_fill_discrete(type = lancet_palette)

mf1
```

### Main figure 2

```{r main-figure-2 }
# Same as PD but does not include 'global'
hbc_predicted <- preds$age_hbc %>%
        select(location, age_group, predicted = point)

# notification data
mf2_full_data <- clean$raw_hbc %>%
        filter(year == 2020) %>%
        group_by(iso3, age_group) %>%
        summarise(cases = sum(cases), .groups = "drop") %>%
        left_join(hbc_predicted,
                  by = c("iso3" = "location", "age_group")) %>%
        left_join(dd$long$country %>%
                          select(country, iso3) %>% distinct(),
                  by = c("iso3")) %>%
        mutate(color = NA_character_) %>%
        select(location = iso3,
               everything())

mf2_data <- mf2_full_data %>%
  filter(location != "LBR")

# calculate order
hbc_order <- mf2_data %>%
        filter(age_group != "15+ years") %>%
        group_by(country) %>%
        summarise(
                dfr = sum(cases - predicted)
        ) %>%
        arrange(dfr) %>% pull(country)

# create percentage burden text labels
age_totals <- mf2_data %>%
        group_by(age_group) %>%
        summarise(total = sum(cases))

mf2_data <- mf2_data %>%
        left_join(age_totals,
                  by = "age_group") %>%
        mutate(pctext = 100 * cases / total,
               pctext = paste0(sprintf(pctext, fmt = '%#.1f'),'%'))

# add HBC29 data
hb29_df <- mf2_data %>%
        group_by(age_group) %>%
        summarise(cases = sum(cases),
                  predicted = sum(predicted)) %>%
        mutate(country = "HBC29",
               pctext = "100%",
               color = "black")

mf2_data <- bind_rows(mf2_data,
          hb29_df) %>%
        mutate(percent_change = 100 * (cases - predicted) / predicted,
               color = case_when(
                       is.na(color) & percent_change < 0 ~ "#930000",
                       is.na(color) & percent_change > 0 ~ "#00639d",
                       TRUE ~ "black"
               ))

# add order
mf2_data$country <- factor(mf2_data$country,
                           levels = c(rev(hbc_order), "HBC29"),
                           ordered = TRUE)


scl <- 1e2
mf2 <- ggplot(mf2_data,
              aes(
                      x = country,
                      y = percent_change / scl,
                      label = pctext,
                      fill = I(color)
              )) +
        geom_bar(stat = "identity") +
        geom_text(aes(x = country,
                      y = 90 / scl)) +
        facet_wrap(~ age_group) +
        scale_y_continuous(limits = c(-100 / scl, 100 / scl),
                           label = scales::percent) +
        coord_flip() +
        labs(y = "Estimated change in notifications in 2020",
             x = "Country or countries") +
        theme_minimal() +
        theme(
                panel.border = element_rect(colour = "black",
                                            size = 1,
                                            fill = NA)
        )


mf2
```

### Main figure 3

```{r main-figure-3}
stringency <- covidchildtb::stringency_index

stringency_2020 <- stringency %>%
        select(-X, -country_name, iso3 = country_code) %>%
        pivot_longer(col = -iso3,
                     names_to = "dates", values_to = "index") %>%
        filter(str_ends(dates, "2020")) %>%
        group_by(iso3) %>%
        summarise(index = mean(index), .groups = "drop")

hbc_predicted <- preds$age_hbc %>%
        select(location, age_group, predicted = point)

# notification data
mf3_data <- clean$raw_hbc %>%
        filter(year == 2020) %>%
        group_by(iso3, age_group) %>%
        summarise(cases = sum(cases), .groups = "drop") %>%
        left_join(hbc_predicted,
                  by = c("iso3" = "location", "age_group")) %>%
        left_join(dd$long$country %>%
                          select(country, iso3) %>% distinct(),
                  by = c("iso3")) %>%
        filter(iso3 != "LBR") %>%
        select(location = iso3,
               everything(), -country) %>%

        mutate(percent_change = 100 * (cases - predicted) / predicted) %>%
        left_join(stringency_2020, by = c("location" = "iso3"))

mf3 <- mf3_data %>%
        filter(!is.na(index)) %>%
        ggscatter(
                x = "index",
                y = "percent_change",
                add = "reg.line",
                conf.int = TRUE,
                xlim = c(23, 70),
                ylim = c(-100, 50),
                xlab = "Mean COVID stringency index 2020",
                ylab = "Percent change in TB notifications"
        ) +
        geom_text_repel(aes(label = location)) +
        facet_wrap(~ age_group, ncol = 1) +
        stat_cor(
                aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
      label.x = 25, label.y = 45, size = 4
      ) + theme_minimal() + grids()

mf3


```

### Main table 1

```{r main-table-1}

mt1_data <- clean$global %>%
  select(-location) %>%
  mutate(sex = factor(sex,
                      levels = c("m", "f", "All"),
                      labels = c("Male", "Female", "All"),
                      ordered = TRUE)) %>%
  group_by(age_group) %>%
  group_split() %>%
  map(function(x) {
  wide <- pivot_wider(
    x,
    id_cols = year,
    names_from = sex,
    values_from = cases
  ) %>%
          select(year, Male, Female)

  wide$abs_diff <- wide[[grep("Male", names(wide))]] - wide[[grep("Female", names(wide))]]
  wide$ratio <- round(wide[[grep("Male", names(wide))]] / wide[[grep("Female", names(wide))]],
                       3)
  wide
  }) %>%
  purrr::reduce(left_join, by = "year")

mt1 <- flextable(mt1_data) %>%
  set_header_labels(
    year = "Year",
    "year",
    Female.x = "Female",
    Male.x = "Male",
    abs_diff.x = "Difference (n)",
    ratio.x = "M:F ratio",
    Female.y = "Female",
    Male.y = "Male",
    abs_diff.y = "Difference (n)",
    ratio.y = "M:F ratio",
    Female = "Female",
    Male = "Male",
    abs_diff = "Difference (n)",
    ratio = "M:F ratio"
  ) %>%
  add_header_row(
    values = c("", "0-4 years",
               "5-14 years",
               "15+ years"),
    colwidths = c(1, 4, 4, 4)
  ) %>%
  set_table_properties(layout = "autofit", width = 1) %>%
  fontsize(size = 9, part = "all") %>%
  theme_box()
mt1
```

### Main table 2

```{r main-table-2}
age_group_totals <- clean$global %>%
        filter(year == 2020) %>%
        group_by(age_group) %>%
        summarise(actual = sum(cases),
                  .groups = "drop") %>%
        mutate(sex = "all") %>%
        select(sex, age_group, actual)


mt2_data <- clean$global %>%
        filter(year == 2020) %>%
        select(-location, -year, actual = cases) %>%
        bind_rows(age_group_totals) %>%
        arrange(age_group, sex) %>%
        left_join(preds$age_sex %>% select(-sd, predicted = point),
                  by = c("age_group", "sex")) %>%
        mutate(sex = factor(sex, levels = c("m", "f", "all"),
                            labels = c("Male", "Female", "All"),
                            ordered = TRUE),
               predicted_text = glue::glue("{predicted} [{as.integer(lower95)} - {as.integer(upper95)}]")) %>%
        mutate(percent_text = glue::glue("{round((actual - predicted) / predicted * 100, 1)} [{round((actual - lower95) / lower95 * 100, 1)} - {round((actual - upper95) / upper95 * 100, 1)}]")) %>%
        mutate(diff_text = glue::glue("{actual - predicted} [{actual - as.integer(lower95)} - {actual - as.integer(upper95)}]")) %>%
  arrange(age_group, sex)

mt2 <- flextable(mt2_data,
                col_keys = c("age_group", "sex", "actual", "predicted_text",
                             "diff_text", "percent_text")) %>%
        set_header_labels(sex = "Sex", age_group = "", actual = "Observed notifications",
                          predicted_text = "Predicted notifications (N, [95% PI])",
                          diff_text = "Difference (N, [95% PI])", percent_text = "% [95% PI]") %>%
        set_table_properties(layout = "autofit", width = 1) %>%
        merge_v(j = ~ age_group) %>%
        # colformat_double(big.mark = ",", digits = 0) %>%
        fontsize(size = 9, part = "all") %>%
        add_footer_row(top = FALSE,
                       values = c("PI: Prediction Interval"),
                       colwidths = c(6)) %>%
        theme_box()

mt2
```

### Supplement figure 1

```{r supp-figure-1, fig.height = 15, fig.width= 10}
age_region_plots <- clean$region %>%
  bind_rows(clean$global) %>%
  group_by(location) %>%
  nest() %>%
  mutate(plots = map2(data, location,
                      function(region, name) {
    ggplot(data = region,
           aes(
             x = year,
             y = cases / 1000,
             group = sex,
             color = sex
           )) +
      geom_line() +
      scale_color_discrete(guide = "none") +
      facet_wrap(. ~ age_group, scales = "free_y") +
      theme_minimal() +
      labs(x = "",
           y = "Cases (thousands)",
           title = name) +
      theme(strip.background = element_blank(),
            strip.text = element_text())

  }))


sf1 <- wrap_plots(age_region_plots$plots, nrow = 7) +
  plot_annotation(title = "Notifications between 2014 and 2020 by age group and WHO Region",
                  caption = "Source: WHO Global TB Programme") +
  scale_color_discrete() +
  theme(legend.position = "bottom")



sf1
```

### Supplement figure 2

```{r supp-fig-2}
# use non-imputed HBC data
md <- calculate_sex_meta_data(data = clean$raw_hbc)

sf2_data <- md$ans %>%
        mutate(iso3 = "SUMMARY") %>%
        select(iso3,
               age_group,
               mid = pred,
               lo = ci.lb,
               hi = ci.ub) %>%
        rbind(md$fans, fill = TRUE) %>%
        mutate(iso3 = factor(iso3,
                             levels = c("SUMMARY", levels(md$fans$iso3))))


sf2 <- ggplot(sf2_data,
       aes(x = iso3, y = mid, ymin = lo, ymax = hi)) +
  geom_point(data = md$M, shape = 1, alpha = 0.5) +
  geom_errorbar(width = 0, col = "blue") +
        geom_point(col = "blue") +
  scale_x_discrete(limits = c("SUMMARY", levels(md$fans$iso3))) +
  scale_y_continuous(label = scales::percent,
                     limits = c(0.25, 0.75)) +
  facet_wrap(~age_group) +
        coord_flip() +
  theme_minimal() +
  geom_hline(yintercept = 0.5, col = 2) +
  labs(y = 'Proportion of notifications from males',
       x = "Country code & summary")

sf2



# create manuscript text
ratios <- md$ans %>%
        mutate(across(.cols = where(is.numeric), ~ round(.x * 100, 1)))

strings <- lapply(ratios$age_group,
       function(age_group) {
               x <- ratios %>%
                       filter(age_group == {{age_group}})
               glue::glue("{x$pred[1]}% (95%CI: {x$ci.lb[1]}% to {x$ci.ub[1]}%)")
       }) %>% set_names(ratios$age_group)

```

#### Text for manuscript

The meta-analytic summary proportions of notifications from males were: `r strings[["0-4 years"]]` for ages 0-4 years; `r strings[["5-14 years"]]` for ages 5-14 years; and `r strings[["15+ years"]]`for ages 15+ years

### Supplement figure 3

```{r supp-figure-3}
countries_u5_change <- dd$long$country %>%
        filter(year %in% c(2019, 2020)) %>%
        # add male and female case together
        pivot_wider(names_from = sex, values_from = cases) %>%
        mutate(cases = m + f) %>%
        group_by(iso3, age_group) %>%
        mutate(percent_change = (cases - lag(cases, 1)) / lag(cases, 1) * 100) %>%
        filter(year == 2020) %>%
        filter(!is.na(percent_change)) %>%
        select(-m, -f) %>%
        mutate(percent_change = if_else(is.finite(percent_change),
                                        percent_change, NA_real_)) %>%
        filter(age_group == "04") %>%
        ungroup() %>%
        select(iso3, percent_change) %>%
        mutate(sqrt_change = sign(percent_change) * sqrt(abs(percent_change)))

countries_u5_change

map_df <- sp::merge(world, countries_u5_change, by = "iso3", all.x = TRUE)

map_sf <- st_as_sf(map_df)

sf3 <- ggplot(data = map_sf,
       aes(fill = sqrt_change)) +
        geom_sf() +
        scale_fill_gradient2(name = "notification change\n(square root of %)",
                             na.value = "grey") +
        theme_minimal() +
  theme(
    legend.position = c(.1, .45),
    legend.title.align = 0.5,
    axis.text.x = element_blank(),
    axis.text.y = element_blank()
  )

sf3



```

### Supplement figure 4

```{r supp-fig-4}

diff_data <- clean$region %>%
        group_by(location, age_group, year) %>%
        summarise(cases = sum(cases),
                  .groups = "drop") %>%
        group_by(location, age_group) %>%
        mutate(diff = cases - lag(cases),
               percent = round(cases / lag(cases) * 100, 1)) %>%
        filter(!is.na(diff))

global_diff_data <- clean$region %>%
        group_by(age_group, year) %>%
        summarise(cases = sum(cases),
                  .groups = "drop") %>%
        group_by(age_group) %>%
        mutate(diff = cases - lag(cases),
               percent = round(cases / lag(cases) * 100, 1),
               location = "Global") %>%
        filter(!is.na(diff))

sf4_data <- bind_rows(global_diff_data,
                      diff_data) %>%
        mutate(location = factor(location,
                                 levels = c("Global",
                                            unique(diff_data$location))),
               alpha = if_else(location == "Global",
                               1, 0.3))

sf4 <- ggplot(sf4_data,
                    aes(x = year,
                        y = percent,
                        color = location,
                        group = location,
                        alpha = alpha)) +
        geom_line() +
        geom_point(color = "black", size = 0.7) +
        geom_hline(yintercept = 100, alpha = 0.4) +
        theme_minimal() +
        coord_cartesian(ylim = c(50, 150)) +
        facet_grid(age_group ~ .) +
        scale_alpha_continuous(guide = "none") +
        scale_color_discrete(type = c("#ff0000", lancet_palette)) +

        labs(y = "Percent change on previous year",
             x = "Year",
             color = "WHO Region")

sf4


```

### Supplement table 1

```{r supp_table-1}

base <- clean$region %>%
  filter(year >= 2019) %>%
  group_by(age_group, sex, year) %>%
  summarise(cases = sum(cases),
            .groups = "drop")

st1_data <- clean$region %>%
  filter(year >= 2019) %>%
  group_by(age_group, year) %>%
  summarise(cases = sum(cases),
            .groups = "drop") %>%
  mutate(sex = "All") %>%
  bind_rows(base) %>%
  pivot_wider(names_from = "year",
              names_prefix = "y_",
              values_from = cases) %>%
   mutate(abs_diff = y_2020 - y_2019,
         perc = round((y_2020 - y_2019) / y_2019 * 100, 2)) %>%
  mutate(sex = factor(sex,
                      levels = c("m", "f", "All"),
                      labels = c("Male", "Female", "All"),
                      ordered = TRUE)) %>%
  arrange(age_group, sex)

st1 <- flextable(st1_data,
                 col_keys = c("age_group", "sex", "y_2019",
                              "y_2020", "abs_diff", "perc")) %>%
  set_header_labels(
    age_group = "",
    sex = "Sex",
    y_2019 = "Observed 2019",
    y_2020 = "Observed 2020",
    abs_diff = "Difference (N)",
    perc = "Percent difference (%)"
  ) %>%
  set_table_properties(layout = "autofit", width = 1) %>%
  merge_v(j = ~ age_group) %>%
  fontsize(size = 9, part = "all") %>%
  theme_box()

st1
```

### Supplement table 2

```{r supp-table-2}

st2_all_regions <- clean$global %>%
  filter(year %in% c(2019, 2020)) %>%
  group_by(age_group, year) %>%
  summarise(cases = sum(cases),
            location = "All",
            .groups = "drop") %>%
  select(location, year, age_group, cases)

st2_data <- clean$region %>%
  filter(year >= 2019) %>%
  group_by(location, year, age_group) %>%
  summarise(cases = sum(cases),
            .groups = "drop") %>%
  bind_rows(st2_all_regions) %>%
  pivot_wider(names_from = "year", values_from = cases,
              names_prefix = "y_" ) %>%
  mutate(abs_diff = y_2020 - y_2019,
         perc = round((y_2020 - y_2019) / y_2019 * 100, 2)) %>%
  mutate(location = factor(location,
                           levels = c(const$who_regions, "All"),
                           ordered = TRUE)) %>%
  arrange(age_group, location)


st2 <- flextable(st2_data,
                 col_keys = c("age_group", "location", "y_2019",
                              "y_2020", "abs_diff", "perc")) %>%
  set_header_labels(
    age_group = "",
    location = "Region",
    y_2019 = "Observed 2019",
    y_2020 = "Observed 2020",
    abs_diff = "Difference (N)",
    perc = "Percent difference (%)"
  ) %>%
  set_table_properties(layout = "autofit", width = 1) %>%
  merge_v(j = ~ age_group) %>%
  fontsize(size = 9, part = "all") %>%
  theme_box()

st2
```

### Supplement table 3

```{r supplement-table-3}
st3_all_regions <- clean$global %>%
  filter(year == 2019) %>%
  group_by(age_group, sex) %>%
  mutate(sex = ifelse(sex == "m", "Male", "Female")) %>%
  summarise(cases = sum(cases),
            location = "All",
            .groups = "drop") %>%
  select(location, sex, age_group, cases)


st3_data <- clean$region %>%
  filter(year == 2019) %>%
  group_by(location, sex, age_group) %>%
  summarise(cases = sum(cases),
            .groups = "drop") %>%
  mutate(sex = factor(sex,
                      levels = c("m", "f", "All"),
                      labels = c("Male", "Female", "All"),
                      ordered = TRUE)) %>%
  bind_rows(st3_all_regions) %>%
  pivot_wider(names_from = "sex", values_from = cases) %>%
  mutate(abs_diff = Male - Female,
         ratio = round(Male / Female, 3)) %>%
  mutate(location = factor(location,
                           levels = c(const$who_regions, "All"),
                           ordered = TRUE)) %>%
  arrange(age_group, location)


st3 <- flextable(st3_data,
                 col_keys = c("age_group", "location", "Male",
                              "Female", "abs_diff", "ratio")) %>%
  set_header_labels(
    age_group = "",
    location = "Region",
    abs_diff = "Difference (N)",
    ratio = "M:F ratio"
  ) %>%
  set_table_properties(layout = "autofit", width = 1) %>%
  merge_v(j = ~ age_group) %>%
  fontsize(size = 9, part = "all") %>%
  theme_box()

st3
```

### Supplement table 4

```{r supp-table-4}
age_group_totals <- clean$global %>%
        filter(year == 2020) %>%
        group_by(age_group) %>%
        summarise(actual = sum(cases),
                  .groups = "drop") %>%
        mutate(location = "All") %>%
        select(location, age_group, actual)

predictions_df <-
  bind_rows(preds$age %>% mutate(location = "All"), preds$age_region) %>%
  select(-sd, predicted = point)


st4_data <- clean$region %>%
        filter(year == 2020) %>%
  group_by(location, age_group) %>%
  summarise(cases = sum(cases),
            .groups = "drop") %>%
        rename(actual = cases) %>%
        bind_rows(age_group_totals) %>%
        arrange(age_group, location) %>%
        left_join(predictions_df,
                  by = c("age_group", "location")) %>%
        mutate(predicted_text = glue::glue("{predicted} [{as.integer(lower95)} - {as.integer(upper95)}]")) %>%
        mutate(percent_text = glue::glue("{round((actual - predicted) / predicted * 100, 1)} [{round((actual - lower95) / lower95 * 100, 1)} - {round((actual - upper95) / upper95 * 100, 1)}]")) %>%
        mutate(diff_text = glue::glue("{actual - predicted} [{actual - as.integer(lower95)} - {actual - as.integer(upper95)}]")) %>%
  mutate(location = factor(location,
                           levels = c(const$who_regions, "All"),
                           ordered = TRUE)) %>%
  arrange(age_group, location)

st4 <- flextable(st4_data,
                col_keys = c("age_group", "location", "actual", "predicted_text",
                             "diff_text", "percent_text")) %>%
        set_header_labels(location = "WHO Region", age_group = "", actual = "Observed notifications",
                          predicted_text = "Predicted notifications (N, [95% PI])",
                          diff_text = "Difference (N, [95% PI])", percent_text = "% [95% PI]") %>%
        set_table_properties(layout = "autofit", width = 1) %>%
        merge_v(j = ~ age_group) %>%
        add_footer_row(top = FALSE,
                       values = c("PI: Prediction Interval"),
                       colwidths = c(6)) %>%
        fontsize(size = 9, part = "all") %>%
        theme_box()

st4
```

### Supplement table 8

```{r supp-table-8}

# format observed
global_totals <- clean$global %>%
        filter(year == 2020) %>%
        mutate(location = "All") %>%
        select(location, age_group, sex,
               actual = cases)

observed_df <- clean$region %>%
        filter(year == 2020) %>%
  group_by(location, age_group, sex) %>%
  summarise(cases = sum(cases),
            .groups = "drop") %>%
        rename(actual = cases) %>%
        bind_rows(global_totals) %>%
        arrange(age_group, location)

stopifnot(nrow(observed_df) == 42)

# predictions
predicted_df <- bind_rows(preds$age_sex_region %>%
  ungroup() %>%
  select(-type),

preds$age_sex %>%
  filter(sex != "all") %>%
  mutate(location = "All") %>%
  select(location, everything(), -sd))

stopifnot(nrow(predicted_df) == 42)


interim_st8_data <- observed_df %>%
  left_join(predicted_df,
                  by = c("age_group", "location", "sex")) %>%
  mutate(
    diff = point - actual,
    diff_lower95 = lower95 - actual,
    diff_upper95 = upper95 - actual
  ) %>%
        mutate(predicted_text = glue::glue("{diff} [{as.integer(diff_lower95)} - {as.integer(diff_upper95)}]")) %>%
  mutate(location = factor(location,
                           levels = c(const$who_regions, "All"),
                           ordered = TRUE)) %>%
  arrange(age_group, location)

ratio_df <- interim_st8_data %>%
  select(location, age_group, sex, diff) %>%
  tidyr::pivot_wider(names_from = "sex", values_from = "diff") %>%
  mutate(mf_ratio = round(m / f, 3)) %>%
  select(-m, -f)

st8_data <- interim_st8_data %>%
  select(-c(
    "actual",
    "point",
    "lower95",
    "upper95",
    "diff_lower95",
    "diff_upper95",
    "diff"
  )) %>%
  tidyr::pivot_wider(names_from = "sex", values_from = "predicted_text") %>%
  left_join(ratio_df,
            by = c("location", "age_group"))

st8 <- flextable(st8_data,
                col_keys = c("age_group", "location", "m", "f", "mf_ratio")) %>%
        set_header_labels(location = "WHO Region", age_group = "",
                          m = "Difference in males (N, [95% PI]) ",
                          f = "Difference in females (N, [95% PI])",
                          mf_ratio = "M:F difference ratio") %>%
        set_table_properties(layout = "autofit", width = 1) %>%
        merge_v(j = ~ age_group) %>%
        add_footer_row(top = FALSE,
                       values = c("PI: Prediction Interval"),
                       colwidths = c(5)) %>%
        fontsize(size = 9, part = "all") %>%
        theme_box()

st8
```

### Supplement table 9

```{r supp-table-9}

st9_data <- mf2_full_data %>%
  select(country, age_group, cases, predicted) %>%
  pivot_wider(names_from = "age_group",
              values_from = c(cases, predicted)) %>%
  select(country,
         cases_0 = `cases_0-4 years`,
         pred_0 = `predicted_0-4 years`,
         cases_5 = `cases_5-14 years`,
         pred_5 = `predicted_5-14 years`,
         cases_15 = `cases_15+ years`,
         pred_15 = `predicted_15+ years`,)


st9 <- st9_data %>%
  flextable() %>%
  set_header_labels(
    country = "Country",
    cases_0 = "Observed 2020",
    pred_0 = "Predicted 2020",
    cases_5 = "Observed 2020",
    pred_5 = "Predicted 2020",
    cases_15 = "Observed 2020",
    pred_15 = "Predicted 2020"
  ) %>%
  add_header_row(top = TRUE,
                 values = c("", levels(mf2_data$age_group)),
                 colwidths = c(1, 2, 2, 2))

st9
```

### Alternative supplement table 8

```{r}
# age groups, WHO regions & global, sexes & all

# prepare predicted
predictions_df <- preds$age_sex_region %>%
  ungroup() %>%
  select(-type) %>%
  bind_rows(preds$age_region %>% select(-sd) %>% mutate(sex = "All")) %>%
  bind_rows(preds$age_sex %>% select(-sd) %>% mutate(location = "All",
                                                     sex = if_else(sex == "all", "All", sex))) %>%
  arrange(location, age_group)



# observed data
# aggregate both sexes
base_observed <- clean$region %>%
  filter(year == 2020) %>%
  bind_rows(clean$global %>% filter(year == 2020) %>%
              mutate(location = "All"))

observed_df <- base_observed %>%
  group_by(location, age_group) %>%
  summarise(cases = sum(cases),
            .groups = "drop") %>%
  mutate(sex = "All") %>%
  bind_rows(base_observed %>% select(-year))

# merge predictions with observed
a_st8_data <- observed_df %>%
  left_join(predictions_df,
            by = c("location", "age_group", "sex")) %>%
  mutate(sex = factor(sex, levels = c("m", "f", "All"),
                            labels = c("Male", "Female", "All"),
                            ordered = TRUE)) %>%
        mutate(percent_text = glue::glue("{round((cases - point) / point * 100, 1)} [{round((cases - lower95) / lower95 * 100, 1)} - {round((cases - upper95) / upper95 * 100, 1)}]")) %>%
  select(age_group, location, sex, percent_text) %>%
  tidyr::pivot_wider(names_from = "sex", values_from = "percent_text") %>%
  mutate(location = factor(location,
                           levels = c(const$who_regions, "All"),
                           ordered = TRUE)) %>%
  arrange(age_group, location)

a_st8_data

a_st8 <- flextable(a_st8_data,
                col_keys = c("age_group", "location", "Male", "Female", "All")) %>%
        set_header_labels(location = "WHO Region", age_group = "",
                          Male = "Percentage difference in males (%, [95% PI]) ",
                          Female = "Percentage difference in females (%, [95% PI])",
                          All = "Percentage difference in all (%, [95% PI])") %>%
        set_table_properties(layout = "autofit", width = 1) %>%
        merge_v(j = ~ age_group) %>%
        add_footer_row(top = FALSE,
                       values = c("PI: Prediction Interval"),
                       colwidths = c(5)) %>%
        fontsize(size = 9, part = "all") %>%
        theme_box()

a_st8
```

```{r save-plots, eval = params$save_artefacts}

ggsave(paste0(output_dir, "/main_arima_plots.", params$plot_type),
       mf1,
       width = 15,
       height = 8)

ggsave(paste0(output_dir, "/main_figure_2.", params$plot_type),
       mf2,
       width = 15,
       height = 8, bg = "white")

ggsave(paste0(output_dir, "/main_figure_3.", params$plot_type),
       mf3,
       width = 7,
       height = 12,
       bg = "white")

ggsave(paste0(output_dir, "/supplement1_age_sex_region_notifcations.",
              params$plot_type),
       plot = sf1,
       width = 10,
       height = 15,
       bg = "white")

ggsave(sf2,
        file = paste0(output_dir, "/supplement2_forest.", params$plot_type),
       w = 10,
       h = 7,
       bg = "white")

ggsave(sf3,
        file = paste0(output_dir, "/supplement3_map.", params$plot_type),
       w = 10,
       h = 7,
       bg = "white")

ggsave(paste0(output_dir, "/supplement4_age_region_annual_change.",
              params$plot_type),
       plot = sf4,
       bg = "white")

```

```{r save-tables, eval=params$save_artefacts}

ps_landscape <- prop_section(page_size = page_size(orient = "landscape"))
ps_portrait <- prop_section(page_size = page_size(orient = "portrait"))

tables_doc <- read_docx() %>%
  body_add_par("Main Table 1", style = "heading 1") %>%
  body_add_flextable(value = mt1) %>%
  body_add_break() %>%
  body_add_par("Main Table 2", style = "heading 1") %>%
  body_add_flextable(value = mt2) %>%
  body_add_break() %>%
  body_add_par("Supplement Table 1", style = "heading 1") %>%
  body_add_flextable(value = st1) %>%
  body_add_break() %>%
  body_add_par("Supplement Table 2", style = "heading 1") %>%
  body_add_flextable(value = st2) %>%
  body_end_block_section(value = block_section(property = ps_portrait)) %>%
  body_add_par("Supplement Table 3", style = "heading 1") %>%
  body_add_flextable(value = st3) %>%
  body_end_block_section(value = block_section(property = ps_landscape)) %>%
  body_add_par("Supplement Table 4", style = "heading 1") %>%
  body_add_flextable(value = st4) %>%
  body_end_block_section(value = block_section(property = ps_landscape)) %>%
  body_add_par("Supplement Table 8", style = "heading 1") %>%
  body_add_flextable(value = st8) %>%
  body_add_par("Alternative supplement Table 8", style = "heading 1") %>%
  body_add_flextable(value = a_st8) %>%
  body_add_par("Supplement Table 9", style = "heading 1") %>%
  body_add_flextable(value = st9)

print(tables_doc, target = paste0(output_dir, "/tables.docx"))

```
